Self-tunable approximated explicit MPC: Heat exchanger implementation and analysys

### 2023-01-27 paper/elsarticle/main.tex
Review: Sec.2.3

%% TODO: UNCOMMENT THIS SECTION TO COMPILE ALSO INITIAL PART OF THE PAPAER
%
%\begin{frontmatter}
%
%%% Title, authors and addresses
%
%%% use the tnoteref command within \title for footnotes;
%%% use the tnotetext command for theassociated footnote;
%%% use the fnref command within \author or \address for footnotes;
%%% use the fntext command for theassociated footnote;
%%% use the corref command within \author for corresponding author footnotes;
%%% use the cortext command for theassociated footnote;
%%% use the ead command for the email address,
%%% and the form \ead[url] for the home page:
%%% \title{Title\tnoteref{label1}}
%%% \tnotetext[label1]{}
%%% \author{Name\corref{cor1}\fnref{label2}}
%%% \ead{email address}
%%% \ead[url]{home page}
%%% \fntext[label2]{}
%%% \cortext[cor1]{}
%%% \affiliation{organization={},
%%%             addressline={},
%%%             city={},
%%%             postcode={},
%%%             state={},
%%%             country={}}
%%% \fntext[label3]{}
%
%\title{Self-tunable approximated explicit MPC: Heat exchanger implementation and analysys}
%%\title{Self-tunable approximated explicit MPC: Implementation on a heat exchanger with a asymmetric behavior}
%
%%% use optional labels to link authors explicitly to addresses:
%%% \author[label1,label2]{}
%%% \affiliation[label1]{organization={},
%%%             addressline={},
%%%             city={},
%%%             postcode={},
%%%             state={},
%%%             country={}}
%%%
%%% \affiliation[label2]{organization={},
%%%             addressline={},
%%%             city={},
%%%             postcode={},
%%%             state={},
%%%             country={}}
%
%\author{Lenka Galčíková}
%\author{Juraj Oravec}
%\author{Petronela Belková}
%
%\affiliation{organization={Slovak University of Technology in Bratislava, Faculty of Chemical and Food Technology, Institute of Information Engineering, Automation, and Mathematics,},%Department and Organization
%            addressline={Radlinského 9}, 
%            city={Bratislava},
%            postcode={81237}, 
%%            state={Slovakia},
%            country={Slovakia}
%}
%
%\begin{abstract}
%%max 250 words!
\textcolor{red}{TODO: Please, checkthe  word limit.}

@@

\textcolor{red}{[TODO: Please, clearly point out the main contributions]}

@@

\textcolor{red}{[TODO: Please, write a paper structure]}

@@

\textcolor{red}{[TODO: Please, fix.]}

@@

\subsection{Self-tunable explicit model predictive control}
\label{sec:self_tunable}
The advantage of tunable controller brings a question of how to design the logic of setting the tuning parameter $\rho$. In this section, the idea of online self-tuning is summarized~\cite{self_tunable}. The concept of self-tuning provides the possibility to adjust the aggressiveness of the controller without the necessity to intervene and tune the penalty matrices during control.

The need for real-time controller tuning often arises from tracking a time-varying reference. The work~\cite{self_tunable} focuses on adjusting the penalty matrix when the reference value is changed. The further the reference value is from the steady state, the more aggressive controller is tuned. The idea behind the suggested scaling liest in compensation of the nonlinear behavior of the system.

The procedure of tuning the controller is based on evaluating the difference between the reference and the steady state, and using this deviation to scale the value of control action. First, the maximal absolute value of reference which can be set during control is defined. This value can be determined based on the general knowledge of future reference values. Another suggestion is to set the maximal deviation $d_{\max}$ simply based on the constraints on system outputs: 
\begin{eqnarray}
	\label{eq:d_max}
	d_{\max} = \max(\vert y_{\min} \vert, y_{\max}),
\end{eqnarray}

where $y_{\min}$ and $y_{\max}$ are respectively lower and upper bound on the output variable. Using the information about the maximal possible deviation $d_{\max}$, the tuning parameter $\rho$ can be calculated as the ratio between the current reference value and the maximal deviation:

\begin{eqnarray}
	\label{eq:rho_self_tunable}
	\rho = \frac{\vert y_{\mathrm{ref}} \vert}{d_{\max}}.
\end{eqnarray}

Based on Eq.~\eqref{eq:rho_self_tunable}, the property $0 \le \rho \le 1$ holds, as $\vert y_{\mathrm{ref}} \vert \le d_{\max}$. As a result, the parameter $\rho$ represents a way how to normalize the deviation from steady state and is exploited to scale the penalty matrix or at the end of the day, the control action.

When considering tuning the control action based on Eq.~\eqref{eq:tunable_u}, a higher value of tuning parameter $\rho$ leads to approaching the upper boundary controller and vice versa. When tuning e.g. the matrix $Q_\mathrm{y}$ penalizing the control error, a higher ratio $\rho$ would lead to more aggressive control actions. When controlling to the reference values closer to the steady state, the parameter $\rho$ decreases and the control becomes sluggish.

@@

Note, the relations \eqref{eq:rho_self_tunable} and \eqref{eq:rho_max} operate with the absolute value of the reference. It is not taken into account, whether the reference change was positive or negative, i.e., whether it changed upwards or downwards. As many plants have asymmetric behavior, the positivity or negativity of the reference change could be considered in the controller tuning to improve the control performance.

@@

\subsection*{Abbreviations}
\begin{center}
	\begin{tabular}{ l l }
		MPC & model predictive control \\
		ISE & integral squared error \\
		LTI & linear time-invariant \\
	\end{tabular}
\end{center}

>>>>

%% TODO: UNCOMMENT THIS SECTION TO COMPILE ALSO INITIAL PART OF THE PAPAER
%
%\begin{frontmatter}
%
%%% Title, authors and addresses
%
%%% use the tnoteref command within \title for footnotes;
%%% use the tnotetext command for theassociated footnote;
%%% use the fnref command within \author or \address for footnotes;
%%% use the fntext command for theassociated footnote;
%%% use the corref command within \author for corresponding author footnotes;
%%% use the cortext command for theassociated footnote;
%%% use the ead command for the email address,
%%% and the form \ead[url] for the home page:
%%% \title{Title\tnoteref{label1}}
%%% \tnotetext[label1]{}
%%% \author{Name\corref{cor1}\fnref{label2}}
%%% \ead{email address}
%%% \ead[url]{home page}
%%% \fntext[label2]{}
%%% \cortext[cor1]{}
%%% \affiliation{organization={},
%%%             addressline={},
%%%             city={},
%%%             postcode={},
%%%             state={},
%%%             country={}}
%%% \fntext[label3]{}
%
%\title{Self-tunable approximated explicit MPC: Heat exchanger implementation and analysys}
%%\title{Self-tunable approximated explicit MPC: Implementation on a heat exchanger with a asymmetric behavior}
%
%%% use optional labels to link authors explicitly to addresses:
%%% \author[label1,label2]{}
%%% \affiliation[label1]{organization={},
%%%             addressline={},
%%%             city={},
%%%             postcode={},
%%%             state={},
%%%             country={}}
%%%
%%% \affiliation[label2]{organization={},
%%%             addressline={},
%%%             city={},
%%%             postcode={},
%%%             state={},
%%%             country={}}
%
%\author{Lenka Galčíková}
%\author{Juraj Oravec}
%\author{Petronela Belková}
%
%\affiliation{organization={Slovak University of Technology in Bratislava, Faculty of Chemical and Food Technology, Institute of Information Engineering, Automation, and Mathematics,},%Department and Organization
%            addressline={Radlinského 9}, 
%            city={Bratislava},
%            postcode={81237}, 
%%            state={Slovakia},
%            country={Slovakia}
%}
%
%\begin{abstract}
%%max 250 words!

\textcolor{red}{[ TODO: Please, prevent using ``When...''-formulation to often. ]}

\textcolor{red}{[ TODO: Please, checkthe  word limit. ]}

@@

\textcolor{red}{[ TODO: Please, clearly point out the main contributions. ]}

@@

\textcolor{red}{[ TODO: Please, write a paper structure. ]}

@@

\textcolor{red}{[ TODO: Please, fix. ]}

@@

\subsection{Self-tunable explicit model predictive control}
\label{sec:self_tunable}

\textcolor{red}{[ TODO: How about ``upgrading'' this subsection into the Section? ]}

The advantage of tunable controller brings a question of how to design the logic of setting the tuning parameter $\rho$. In this section, the idea of online self-tuning is summarized~\cite{self_tunable}. The concept of self-tuning provides the possibility to adjust the aggressiveness of the controller without the necessity to intervene and tune the penalty matrices during control.

The need for real-time controller tuning often arises from tracking a time-varying piece-wise constant (PWC) reference. The work~\cite{self_tunable} focuses on adjusting the penalty matrix when the reference value is changed. The further the reference value is from the steady state, the more aggressive controller is tuned. The idea behind the suggested scaling liest in compensation of the nonlinear behavior of the system.

Consider a single-input and single-output (SISO) system or a multiple-inputs and multiple-outputs (MIMO) system with a completely decoupled pairs of the control inputs and the system outputs. 
Then, the procedure of tuning the controller is based on evaluating the different operating points between the current value of the reference and the system steady-state value. This deviation is considered to scale the value of control action. First, the maximal admissible absolute value of reference is defined. Analogous to the reference trajectory preview concept of MPC design, this value can be determined based on the general knowledge of the expected future reference values. Another suggestion is to fit the maximal deviation $d_{\max}$ based on the constraints on system outputs: 
\begin{eqnarray}
	\label{eq:d_max}
	d_{\max} = \max(\vert y_{\min} \vert, y_{\max}),
\end{eqnarray}
where $y_{\min}$ and $y_{\max}$ are respectively lower and upper bound on the output variable in the deviation form, ie., zero corresponds to the system steady-state value. Using the information about the maximal possible deviation $d_{\max}$, the tuning parameter $\rho$ can be calculated as the ratio between the current reference value and the maximal deviation:  
\begin{eqnarray}
	\label{eq:rho_self_tunable}
	\rho = \frac{\vert y_{\mathrm{ref}} \vert}{d_{\max}}.
\end{eqnarray}

Based on Eq.~\eqref{eq:rho_self_tunable}, the property $0 \le \rho \le 1$ holds, as $\vert y_{\mathrm{ref}} \vert \le d_{\max}$. As a consequence, the parameter $\rho$ represents a way how to normalize the deviation from the steady-state value and is exploited to scale the penalty matrix or, implicitly, to scale the aggressiveness of the control action.

When considering tuning the control action based on Eq.~\eqref{eq:tunable_u}, a higher value of tuning parameter $\rho$ leads to approaching the upper boundary controller and vice versa. When tuning, e.g., the matrix $Q_\mathrm{y}$ penalizing the control error, a higher ratio $\rho$ would lead to more aggressive control actions. When operating with the reference value close to the system steady-state value, the parameter $\rho$ decreases and the control profiles becomes sluggish.

@@

Note, the relations \eqref{eq:rho_self_tunable} and \eqref{eq:rho_max} operate with the absolute value of the reference. It is not taken into account, whether the reference change was positive or negative with respect to the system steady-state value placed in the origin, i.e., whether it changed upwards or downwards. As many plants have asymmetric behavior, the positivity or negativity of the reference change could be considered in the controller tuning to improve the control performance.

@@

\subsection*{Abbreviations}
\begin{center}
	\begin{tabular}{ l l }
		MIMO & multiple-inputs and multiple-outputs (system) \\
		MPC  & model predictive control \\
		ISE  & integral squared error \\
		LTI  & linear time-invariant (system) \\
		mp-QP& multi-parametric quadratic programming (problem) \\
		PWA  & piece-wise affine (function) \\
		PWC  & piece-wise constant (function) \\
		QP   & quadratic programming (problem) \\
		SISO & single-input and single-output (system) \\
	\end{tabular}
\end{center}
